Bioconductor workflow for single-cell RNA sequencing: Normalization, dimensionality reduction, clustering, and lineage inference

### 2017-07-01 workflow/workflow.Rmd
Edits to text and figure captions.

keywords: single-cell, single-cell genomics, RNA-seq, gene expression, zero inflation, normalization, dimensionality reduction, clustering, lineage inference, differential expression, workflow, R, Bioconductor

@@

[FP: we need to have only 8 keywords, which ones do you prefer?]

@@

```{r boxplotNorm, fig.cap="Boxplots of normalized expression measures, color-coded by batch."}
norm_order <- norm[, order(as.numeric(batch))]
col_order <- col_batch[batch[order(as.numeric(batch))]]
boxplot(norm_order, col = col_order, staplewex = 0, outline = 0,
        border = col_order, xaxt = 'n')
abline(h=0)
```

@@

```{r pcanorm, fig.cap="PCA of normalized expression measures, where each point represents a cell. In the left panel, cells are color-coded by batch. In the right panel, cells are color-coded by original published clusters."}
pca <- prcomp(t(norm))
par(mfrow = c(1,2))
plot(pca$x, col = col_batch[batch], pch = 20, main = '')
plot(pca$x, col = col_clus[as.character(publishedClusters)], pch = 20, main = '')
```

@@

```{r mdsW, fig.cap="MDS of the low-dimensional matrix W, where each point represents a cell and cells are color-coded by original published clusters."}
W <- colData(se)[, grepl('^W', colnames(colData(se)))]
W <- as.matrix(W)
d <- dist(W)
fit <- cmdscale(d, eig = TRUE, k = 2)
plot(fit$points, col = col_clus[as.character(publishedClusters)], main = '',
     pch = 20, xlab = 'Component 1', ylab = 'Component 2')
legend(x = 'bottomright', legend = unique(names(col_clus)), cex = .5,
       fill = unique(col_clus), title = 'Sample')
```

@@

```{r examineCombineMany, fig.cap="Clusters found using the function RSEC from the clusterExperiment package."}
plotClusters(ceObj, colPalette = c(bigPalette, rainbow(199)))
```

@@

```{r barplotOurs, fig.cap="Barplot of number of cells per cluster found using our workflow."}
plotBarplot(ceObj) 
```

@@

```{r addPublishedClusters, fig.cap="Barplot of number of cells per cluster in original published clusters."}
ceObj<-addClusters(ceObj, colData(ceObj)$publishedClusters, 
                   clusterLabel = "publishedClusters")
plotBarplot(ceObj,whichClusters=c("mergeClusters","publishedClusters")) 
```

@@

```{r comparePublished, fig.cap="Comparison between the clusters found using our workflow and original published clusters."}
plotClusters(ceObj, whichClusters=c("mergeClusters","combineMany","publishedClusters"))
```

@@

```{r curves, fig.cap="Cells color-coded by cluster in a 4-dimensional space with smooth curves representing each inferred lineage."}
pairs(lineages, type='curves', pal[primaryClusterNamed(ceObj)])
```

```{r tree, fig.cap="Cells color-coded by cluster in a 4-dimensional space with connecting lines between cluster centers representing the inferred structure."}
pairs(lineages, type='lineages', col = pal[primaryClusterNamed(ceObj)])
```

>>>>

keywords: single-cell, RNA-seq, normalization, dimensionality reduction, clustering, lineage inference, differential expression, workflow

@@

[FP: we need to have only 8 keywords, which ones do you prefer? SD: I selected 8, please check.]

@@

```{r boxplotNorm, fig.cap="Boxplots of normalized expression measures (deviance residuals), color-coded by batch."}
norm_order <- norm[, order(as.numeric(batch))]
col_order <- col_batch[batch[order(as.numeric(batch))]]
boxplot(norm_order, col = col_order, staplewex = 0, outline = 0,
        border = col_order, xaxt = 'n', ylab='Expression measure')
abline(h=0)
```

@@

```{r pcanorm, fig.cap="PCA of normalized expression measures, where each point represents a cell. In the left panel, cells are color-coded by batch. In the right panel, cells are color-coded by original published clustering."}
pca <- prcomp(t(norm))
par(mfrow = c(1,2))
plot(pca$x, col = col_batch[batch], pch = 20, main = '')
plot(pca$x, col = col_clus[as.character(publishedClusters)], pch = 20, main = '')
```

@@

```{r mdsW, fig.cap="MDS of the low-dimensional matrix W, where each point represents a cell and cells are color-coded by original published clustering."}
W <- colData(se)[, grepl('^W', colnames(colData(se)))]
W <- as.matrix(W)
d <- dist(W)
fit <- cmdscale(d, eig = TRUE, k = 2)
plot(fit$points, col = col_clus[as.character(publishedClusters)], main = '',
     pch = 20, xlab = 'Component 1', ylab = 'Component 2')
legend(x = 'bottomright', legend = unique(names(col_clus)), cex = .5,
       fill = unique(col_clus), title = 'Sample')
```

@@

```{r examineCombineMany, fig.cap="Candidate clusterings found using the function RSEC from the clusterExperiment package."}
plotClusters(ceObj, colPalette = c(bigPalette, rainbow(199)))
```

@@

```{r barplotOurs, fig.cap="Barplot of number of cells per cluster, for our workflow's clusterint."}
plotBarplot(ceObj) 
```

@@

```{r addPublishedClusters, fig.cap="Barplot of number of cells per cluster, for original published clustering."}
ceObj<-addClusters(ceObj, colData(ceObj)$publishedClusters, 
                   clusterLabel = "publishedClusters")
plotBarplot(ceObj,whichClusters=c("mergeClusters","publishedClusters")) 
```

@@

```{r comparePublished, fig.cap="Comparison between the clustering found using our workflow and the original published clustering."}
plotClusters(ceObj, whichClusters=c("mergeClusters","combineMany","publishedClusters"))
```

@@

```{r curves, fig.cap="Cells color-coded by cluster in a 4-dimensional MDS space with smooth curves representing each inferred lineage."}
pairs(lineages, type='curves', pal[primaryClusterNamed(ceObj)])
```

```{r tree, fig.cap="Cells color-coded by cluster in a 4-dimensional MDS space with connecting lines between cluster centers representing the inferred structure."}
pairs(lineages, type='lineages', col = pal[primaryClusterNamed(ceObj)])
```
